The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities

### 2023-12-07 r_r/letter_to_reviewers.Rmd
revising letter

We thank the Reviewers and the Editor for excellent, thoughtful, and detailed feedback. While the analyses have changed only modestly, the front end of the paper has changed tremendously, literally rewritten from scratch. I begin this letter by discussing the big changes in the manuscript, which I hope address many of the big concerns from Reviewer 2. Next, I address remaining reviewer concerns one by one.

@@

Reviewer 2 very helpfully pointed out several core flaws with assumptions, practical motivations, scholarly contribution, and practical significance of the manuscript. I view these flaws as distinct yet interconnected. First, Reviewer 2 found the assumptions about "mission" -- public research universities serve social mobility mission while selective private universities serve social reproduction -- to be overly simplistic generalizations that assume a homogeneous mission across very different organizations. Moreover, Reviewer 2 found the assumption that private universities serve the mission of social reproduction to be somewhat off-putting and dismissive. Second, in contrast to the emphasis on an idealized mission, Reviewer 2 encouraged the manuscript to consider a more practical reason for why public research universities would target private high schools, particularly around targeting high-paying students that generate substantial revenue. Third, with respect to scholarly contribution, Reviewer 2 questioned why comparing public and private universities on a behavior associated with private universities makes an important contribution to the privatization literature. We must admit that we agree with this concern. The "privatization" literature has been in decline for more than a decade. So it is not important to contribute to a literature that is no longer relevant. Fourth, Reviewer 2 is concerned that the original manuscript did not clearly state the practical implications of the findings.

@@

We develop a "bridging" scholarly contribution by stating the these two literatures have been isolated from one another, but they are both parts of a broader story about the changing nature of school segregation and the competition for affluent students. As public universities seek affluent out-of-state students, they are attracted to out-of-state private schools. In doing so, public research universities are encroaching on what had primarily been the territory of private universities. Therefore, analyses of visits to private schools explore the extent to which public and private universities are competing for students at similar or distinct sets of schools, revealing insights about the competition for affluent students. Additionally, prior research shows that private school enrollment is whiter than public schools, that private school students have become more affluent in recent decades, and that private school enrollment has grown in the South as a consequence of "white flight" to private schools. "Thus, in recruiting from private schools, universities may be targeting student bodies that are more affluent and racially homogeneous than public schools. We investigate the characteristics – including geographic region and racial composition – of private schools being targeted by university admissions officers because these characteristics have downstream implications for the racial composition of university campuses"

Finally, we have revised the practical significance and implications to address concerns raised by Reviewer 2.

[WRITE THIS TEXT AFTER YOU REVISE THE DISCUSSION SECTION]

@@

**Author response**: We wholeheartedly agree with these concern by Reviewer 2 and we are grateful to Reviewer 2 for raising these issues! As discussed above, the revised manuscript addresses these concerns about the practical benefits of recruiting students from private schools. We have also improved the discussion of implications for research and practice [REVISE AFTER YOU REWRITE THE DISCUSSION SECTION]

@@

**Author response**: REVISE THIS RESPONSE AFTER YOU HAVE WRITTEN THE DISCUSSION SECTION AND REVISED THE INTRODUCTION.

@@

[REVISE THIS TEXT AFTER YOU REVISE THE DISCUSSION SECTION]
Nevertheless, the question Reviewer 2 asks is very important; if off-campus recruiting visits became less prevalent following Covid, then the analyses do yield relevant insights about contemporary/future recruiting behavior by public research universities and selective private universities.

In the revised manuscript, the discussion section substantively describes information about off-campus recruiting events following Covid. In short, off-campus recruiting events are back and are viewed as vital to recruiting efforts. We put this text in the discussion section rather than the literature review because it complements text about recommendations for practice and this text fits lesss well with the revised literature review structure.

TEXT TEXT TEXT This recommendation is very relevant for the original manuscript, but a bit less relevant for the revised manuscript in that the previous lit review of "what do we know about off-campus recruiting" has been replaced by a much broader literature review which focuses on X, Y, Z. With that being said, the reviewer asks a good question here. We refer [IN FOOTNOTE?] the readers to REPORT X, Y, Z FROM RUFFALO NOEL LEVITZ WHICH SAYS BLAH BLAH BLAH. FROM EAB COUNSELOR REPORT Q16, "HOW DO YOU PREFER TO RECEIVE INFORMATION FROM COLLEGES"; FOR BOTH PUBLIC AND PRIVATE SCHOOLS, "HIGH SCHOOL VISITS BY COLLEGE REPRESENTATIVES" WERE THE HIGHEST RATED CHANNEL, ESPECIALLY FOR PRIVATE SCHOOL COUNSELORS. ALSO, VIRTUAL COUNSELOR EVENTS WERE QUITE LOW RATED. IMPLICATION, IN THE PURSUIT OF TRADITIONAL COLLEGE STUDENTS, THIS IS STILL A HAND-TO-HAND BUSINESS. ALSO THERE IS A BIT IN THE MARKETING RECRUITING PRACTICES FOR UNDERGRAD STUDENTS BY RUFFALO.

Cost of recruiting undergraduates report, 2022. Travel was 5th for publics. second for privates. Travel was top area of increased expenditure for publics and second top area of increase for privates, suggesting that they were going to increase travel following return from Covid.

@@

INCLUDE PICTURE OF EGO NETWORK? HAVE TO KEEP THIS PARSIMONIOUS BECAUSE OF PAGE LENGTH CONCERNS

@@

ADD APPENDIX TABLE ABOUT THIS.

@@

**Author response**: Thank you! The results have been revised moderately to reflect analysis of recruiting in neighboring state and to reflect questions from Reviewer 1 (ADD/MODIFY TEXT ONCE YOU HAVE MADE THIS CHANGE)

@@

STATE THE SCHOLARLY CONTRIBUTION THE PRESENT ARTICLE DOES MAKE.

- TO THE NONRES LITERATURE/COMPARE THESE RESULTS TO RESULTS FROM AERJ ARTICLE. SAY THE CONTRIBUTION
- TO THE PRIVATE SCHOOL TO PRIVATE UNIVERSITY LINKAGES, SHOW WHAT ARE THE ORGS THAT UNIVERSITIES ESTABLISH TIES TO (E.G., ALONG RELIGIOUS LINES)
- WHAT IS THE BRIDGING CONTRIBUTION.

@@

**Author response**: ADD TEXT

>>>>

We thank the Reviewers and the Editor for excellent, thoughtful, and detailed feedback. While the analyses have changed only modestly, the front end of the paper has changed tremendously, literally rewritten from scratch. I begin this letter by describing the big changes in the manuscript, which I hope address many of the big concerns from Reviewer 2. Next, I address remaining reviewer concerns one by one.

@@

Reviewer 2 very helpfully pointed out several core flaws with assumptions, practical motivations, scholarly contribution, and practical significance of the manuscript. I view these flaws as distinct yet interconnected. First, Reviewer 2 found the assumptions about "mission" -- public research universities serve social mobility mission while selective private universities serve social reproduction -- to be overly simplistic generalizations that assume a homogeneous mission across very different organizations. Moreover, Reviewer 2 found the assumption that private universities serve the mission of social reproduction to be somewhat off-putting and dismissive. Second, in contrast to the emphasis on an idealized mission, Reviewer 2 encouraged the manuscript to consider a more practical reason for why public research universities would target private high schools, particularly around targeting high-paying students that generate substantial revenue. Third, with respect to scholarly contribution, Reviewer 2 questioned why comparing public and private universities on a behavior associated with private universities makes an important contribution to the privatization literature. We must admit that we agree with this concern. The "privatization" literature has been in decline for more than a decade. It is not important to contribute to a literature that is no longer relevant. Fourth, Reviewer 2 is concerned that the original manuscript did not clearly state the practical implications of the findings.

@@

We develop a "bridging" scholarly contribution by stating the these two literatures have been isolated from one another, but they are both parts of a broader story about the changing nature of school segregation and the competition for affluent students. As public universities seek affluent out-of-state students, they are attracted to out-of-state private schools. In doing so, public research universities are encroaching on what had primarily been the territory of private universities. Therefore, analyses of visits to private schools explore the extent to which public and private universities are competing for students at similar or distinct sets of schools, revealing insights about the competition for affluent students. Additionally, prior research shows that private school enrollment is whiter than public schools, that private school students have become more affluent in recent decades, and that private school enrollment has grown in the South as a consequence of "white flight" to private schools. "Thus, in recruiting from private schools, universities may be targeting student bodies that are more affluent and racially homogeneous than public schools. We investigate the characteristics – including geographic region and racial composition – of private schools being targeted by university admissions officers because these characteristics have downstream implications for the racial composition of university campuses."

The research questions in the revised manuscript are very similar to the research questions from the original manuscript. Two specific changes:

1. First, we switched the order of RQ2 and RQ3. In the revised manuscript, overlap in recruiting network is RQ2 and characteristics of visited high schools is RQ3. Our rationale is that the characteristics of visited private high schools is the most climactic, substantively important set of results, so we wanted those to go last. Additionally, when thinking through the narrative story of the results, we thought it made more sense to identify overlap -- extent of overlap, which universities visiting similar sets of high schools -- and then examine the characteristics of visited high schools.
1. Second, compared to the original manuscript, the revised manuscript research questions are worded in a way that places less emphasis on comparing public research universities to selective private universities. The original manuscript focused on whether public research universities behave like selective private universities. Here, comparison is central. The revised manuscript is based on the idea both public research universities and selective private universities are targeting the same (or similar) sets of predominantly affluent households. Although comparison remains important, the fundamental interest now becomes developing insights about the competition for students who attend private schools and which schools the universities in our sample are targeting.

Finally, in the Discussion section, we have revised the practical significance and implications for colleges and universities to address concerns raised by Reviewer 2. The final paragraph of the revised introduction describes these implications succinctly:

>This manuscript yields practical implications for public and private universities concerned about racial and socioeconomic diversity. Following the Covid-19 Pandemic, off-campus recruiting visits have bounced back because they remain a fundamental means of maintaining relationships with high school guidance counselors [@RN4896; @RN4895; @eab_nacac_counselor_2023]. Prior research finds that that students are sensitive to recruiting interventions, including off-campus recruiting visits [@RN4324; @RN4402]. Therefore, which high schools a university chooses to visit affect the subsequent enrollment composition of the university. In the aftermath of the Supreme Court decision on race conscious admissions, public and private universities are searching for ways to increase racial diversity without running afoul of the law. The @RN4954[p. 9] recommends that universities expand "outreach to high schools with substantial populations of low-income students and students of color." Most public and private universities in our sample visited a disproportionate number of private schools and these private schools tended to enroll mostly white students. We recommend that universities concerned about racial diversity should conduct an internal equity audit of their recruiting practices, for example, comparing the racial and socioeconomic characteristics of schools they visit versus schools they do not visit.

@@

**Author response**: We wholeheartedly agree with these concern by Reviewer 2 and we are grateful to Reviewer 2 for raising these issues! As discussed above, the revised manuscript addresses these concerns about the practical benefits of recruiting students from private schools. We have also improved the discussion of implications for research and practice.

@@

**Author response**: As stated above, the last paragraph of the revised introduction introduces practical implications. The Discussion section discusses these implications in greater detail. We focus on implications for universities rather than implications for policy because the implications for universities are clearer than the implications for policymakers.

@@

Nevertheless, the question Reviewer 2 asks is very important; if off-campus recruiting visits became less prevalent following Covid, then the analyses do yield relevant insights about contemporary/future recruiting behavior by public research universities and selective private universities.

In the revised manuscript, the discussion section substantively describes information about off-campus recruiting events following Covid. In short, off-campus recruiting events are back and are viewed as vital to recruiting efforts. We put this text in the discussion section rather than the literature review because it complements text about recommendations for practice and this text fits less well with the revised literature review structure. Here is the revised text that we placed at the beginning of the "implications for practice" sub-section of the Discussion section.

> Market research suggests that off-campus recruiting visits will remain a vital means of connecting to students and guidance counselors, just as they were prior to the Covid pandemic. @RN4664 found that "travel" was the top-ranked top-ranked marketing/recruiting expenditure item for both public universities and private universities. During the height of Covid, off-campus recruiting visits ceased and were partially replaced by virtual counselor events [@parents_covid_rnl]. Admissions counselors are now traveling again. @RN4896 found that travel was the fifth ranked expenditure item for public universities and the second ranked expenditure item for private universities. When asked about areas of future expenditure growth, travel was the top ranked item for public universities and the second ranked item for private universities [@RN4896], suggesting that universities are increasing travel following the decline of the pandemic.

> The @eab_nacac_counselor_2023 survey of high school counselors asked "how do you prefer to receive information from colleges?" High school visits by college representatives was by far the highest rated "channel" for private school counselors and was also the highest rated channel for public school counselors. By contrast, virtual visits by admissions representatives received low ratings by high school guidance counselors. Finally, @RN4895 asked universities to rank the most effective "outreach strategies for high school counselors." "Calling after visits" was deemed effective for 97% of private universities (the top strategy) and was deemed effective by 88% of public universities. What insights do we glean from these market research findings? In the era of mass marketing and data science, the pursuit of traditional college students from privileged schools and communities is still a hand-to-hand business and off-campus recruiting visits remain a vital means of recruiting these students.

@@

**Author response**: We appreciate this concern and agree with the recommendation. We thought of several tables and figures that could help readers visualize some of the network concepts. However, we made the decision to exclude most of these because the manuscript is already quite long. The revised manuscript does include a visualization of the full two-mode recruiting network between universities and high schools (CRYSTAL STATE FIGURE NUMBER ONCE YOU ADD IT). The manuscript also includes a visualization of the ego network of University of Notre Dame (Figure 1), which is useful in explaining what are ego networks and their relationship with RQ3 (characteristics of high schools that reveive visits)

@@

**Author response**: We agree with this suggestion. The revised manuscript now includes a table (Table 2) that compares the universities in the analysis sample to the relevant populations the analysis sample was drawn from.

@@

**Author response**: Thank you!

@@

As stated above, the revised manuscript attempts to set up the scholarly contribution as bridging two previously disconnected literatures: (1) the literature on nonresident enrollment at public research universities; and (2) the literature on linkages between private schools and selective (private) universities.

In the "implications for scholarship" sub-section of the revised Discussion section, we discuss the contribution of this manuscript to the two respective literatures. Next, we argue that our broader/more important contribution is showing that these literatures are part of the same story and developing new insights about the competition privileged students. Here, I paste text from the revised manuscript:

> Our more important scholarly contribution is showing that these two literatures are both parts of a broader story about the changing nature of competition for students from privileged schools and communities. Selective private universities have been recruiting heavily from private schools for a long time [@RN4642; @RN3519; @RN4640]. By contrast, the pursuit of out-of-state private school students by public research universities is a relatively new phenomenon and represents competition in the traditional hunting grounds of selective private universities.

> Our analyses shed light about the nature of this competition. When making decisions about which private schools to visit, public research universities seem to be mimicking the behavior of particular private universities rather than mimicking the behavior of private universities writ large. Several public universities seem to be competing directly for students targeted by a particular private university. For example, UC-Berkeley visited 27.9% of the schools visited by Northwestern, 28.9% of the schools visited by Tufts, and 27.8% of the schools visited by Emory (Figure \@ref(tab:privu-percent-visited-by-pubu-matrix)). We reason UC-Berkeley is competing directly for students because all four universities are in the same broad echelon of prestige. Other potential examples of direct competition include the University of Georgia targeting schools visited by Emory, SMU, TCU, and Tulane, or the University of Arkansas targeting schools visited by Baylor.

> However, public universities that made the greatest number of private school visits seemed to be targeting lower-achieving students at private schools targeted by particular private universities. For example, the University of Alabama visited 52.2% of the 625 private schools that Notre Dame visited, with most of the overlap at Catholic schools. Given the differences in academic prestige between them, the University of Alabama may not be competing for the same students as Notre Dame. Instead, they seem to be targeting Catholic school students of a similar economic echelon, who have lower academic achievement, and want to attend a football powerhouse. Additionally, CU-Boulder visited 42.4% of schools visited by Northwestern, which could be interpreted as CU-Boulder competing for lower-achieving students of a similar economic and cultural echelon at the schools targeted by Northwestern. Similarly, the University of South Carolina competes heavily at schools targeted by Emory, SMU, TCU, and Tulane.

@@

**Author response**: ADD TEXT ON FRIDAY.

### 2023-12-07 r_r/manuscript_rihe_r_r.Rmd
revising letter

This manuscript yields practical implications for public and private universities concerned about racial and socioeconomic diversity. Following the Covid-19 Pandemic, off-campus recruiting visits have bounced back because they remain the fundamental means of maintaining relationships with high school guidance counselors [CITE]. Prior research shows that that students are sensitive to recruiting interventions, including off-campus recruiting visits. Therefore, which high schools a university chooses to visit will affect the subsequent enrollment composition of the university. In the aftermath of the Supreme Court decision on race conscious admissions, public and private universities are searching for ways to increase racial diversity without running afoul of the law. The U.S. Department of Education recommends that universities expand "outreach to high schools with substantial populations of low-income students and students of color" (CITE, PX X). Most public and private universities in our sample visited a disproportionate number of private schools and these private schools tended to enroll mostly white students. We recommend that universities concerned about racial diversity should conduct an internal equity audit of their recruiting practices, for example, comparing the racial and socioeconomic characteristics of schools they visit versus schools they do not visit.

@@

Degree is a limited measure of centrality in that it only identifies direct connections. By contrast, *k-path centrality* measures the extent to which an actor/vertex is indirectly connected to others, *betweenness centrality* measures the extent to which a vertex is an intermediary between otherwise disconnected groups of actors, and *eigenvector centrality* measures the extent to which a vertex is connected to highly-connected others [@RN4725]. Betweenness centrality tends to be relevant for empirical contexts (e.g., which manager gets credit for an idea) where an actor who occupies a position between groups can benefit by being a broker of information/ideas [@RN975; @RN4648]. By contrast, this chapter is more concerned with whether a particular college tends to have indirect connections with some colleges but not others, and what is the subset of high schools this group of colleges tends to visit. Research question 2, motivated below, explores these ideas.

@@

This study engages with two, previously disconnected literatures. Scholarship on nonresident enrollment observes that many public research universities began growing nonresident enrollment following declines in state appropriations []. Given that nonresident tuition price is often two- to three-times resident price, it is unsurprising that nonresident students tend to be more affluent than resident students, are less likely to be first-generation students, and are less likely to be Black or Latinx [CITE]. Public research universities recruit nonresident students using institutional financial aid packages [CITE] and through recruiting interventions that target earlier stages in the enrollment funnel. Using nearly the same sample of 15 public research universities used in this study, Author (XXXX) found that 12 universities made more out-of-state visits than in-state visits and visits to out-of-state public high schools focused on affluent, predominantly white schools. Public research universities have a financial incentive to target private school students, who tend to come from affluent households able to pay nonresident tuition price without need-based aid. This study extends the literature on nonresident students by analyzing the scale of visits to private schools, the overlap and divergence in recruiting networks between public and private universities, and the characteristics of visited private schools.

We also engage with scholarship on linkages between private schools and private universities, a linkage that sociologists have observed for many decades. Private school students comprise a disproportionate share of enrollment at selective private universities [CITE; AND AUTHOR CALCULATIONS]. Practically, these linkages exist because private schools want to send their students to "good" universities while selective private universities want to enroll students who can pay private university tuition and who have desired academic profile and/or extracurricular activities. However, prior research tends to be focused on elite schools and universities [CITE KAHN; CHETTY; CHARTER+BARTER] and does not systematically analyzed connections between private schools and private universities that are selective, but not elite. This study extends the literature by investigating private school visits for a sample of 14 selective private universities that exhibit variation in geographic location, religious affiliation, and ranking.

Our more important scholarly contribution is showing that these two separate stories are both parts of a broader narrative about the changing nature of competition for students from privileged schools and communities. Selective private universities have been recruiting  heavily from private schools for a long time [@RN4642; @RN3519; @RN4640]. By contrast, the pursuit of out-of-state private school students by public research universities is a relatively new phenomenon and represents competition in the traditional hunting grounds of selective private universities.

@@

However, public universities that made the greatest number of private school visits seemed to be targeting lower-achieving students at private schools targeted by particular private universities. For example, the University of Alabama visited 52.2% of the 625 private schools that Notre Dame visited, with most of the overlap at Catholic schools. Given the differences in academic prestige between them, the University of Alabama may not be competing for the same students as Notre Dame. Instead, they found a niche by targeting Catholic school students of a similar economic echelon, who have lower academic achievement, and want to attend a football powerhouse. Additionally, CU-Boulder visited 42.4% of schools visited by Northwestern, which could be interpreted as CU-Boulder competing for lower-achieving students of a similar economic and cultural echelon at the schools targeted by Northwestern. Similarly, the University of South Carolina competes heavily at schools targeted by Emory, SMU, TCU, and Tulane.

@@

Market research suggests that off-campus recruiting visits will remain a vital means of connecting to students and guidance counselors, just as they were prior to the Covid pandemic. @RN4664 found that "travel" was the top-ranked top-ranked marketing/recruiting expenditure item for both public universities and private universities. During the height of Covid, off-campus recruiting visits ceased and were partially replaced by virtual counselor events [CITE]. Admissions counselors are now traveling again. @RN4896 found that travel was the fifth ranked expenditure item for public universities and the second ranked expenditure item for private universities. When asked about areas of future expenditure growth, travel was the top ranked item for public universities and the second ranked item for private universities @RN4896, suggesting that universities are increasing travel following the decline of the pandemic.

Additionally, the EAB (XXXX) survey of high school counselors asked "how do you prefer to receive information from colleges?" High school visits by college representatives was the highest rated "channel" for both public school counselors and (especially) for private school counselors. By contrast, virtual visits by admissions representatives received low ratings by high school guidance counselors. Finally, @RN4895 asked universities to rank the most effective "outreach strategies for high school counselors." "Calling after visits" was deemed effective for 97% of private universities (the top strategy) and was deemed effective by 88% of public universities. What insights do we glean from these market research findings? In the era of mass marketing and data science, the pursuit of traditional college students from privileged schools and communities is still a hand-to-hand business and off-campus recruiting visits remain a vital means of recruiting these students.

Considering the practical importance of off-campus recruiting visits, what are the practical implications of our research? Prior scholarly and market research [] shows that students are sensitive to recruiting interventions, including recruiting visits. This means that racial and socioeconomic composition of the schools that a university chooses to visit affects the racial and socioeconomic composition of the university. Our research on visits to public high schools [CITE] and this study on visits to private schools finds that our sample of public research universities and selective private universities tend to focus their recruiting efforts on affluent and predominantly white schools and communities.    Some universities may publicly proclaim a commitment to social mobility and privately pursue an intentional recruiting strategy that targets affluent schools and communities. We offer practical guidance for universities that may be unintentionally implementing a recruiting strategy tied to social reproduction.

>>>>

This manuscript yields practical implications for public and private universities concerned about racial and socioeconomic diversity. Following the Covid-19 Pandemic, off-campus recruiting visits have bounced back because they remain a fundamental means of maintaining relationships with high school guidance counselors [@RN4896; @RN4895; @eab_nacac_counselor_2023]. Prior research finds that that students are sensitive to recruiting interventions, including off-campus recruiting visits [@RN4324; @RN4402]. Therefore, which high schools a university chooses to visit affect the subsequent enrollment composition of the university. In the aftermath of the Supreme Court decision on race conscious admissions, public and private universities are searching for ways to increase racial diversity without running afoul of the law. The @RN4954[p. 9] recommends that universities expand "outreach to high schools with substantial populations of low-income students and students of color." Most public and private universities in our sample visited a disproportionate number of private schools and these private schools tended to enroll mostly white students. We recommend that universities concerned about racial diversity should conduct an internal equity audit of their recruiting practices, for example, comparing the racial and socioeconomic characteristics of schools they visit versus schools they do not visit.

@@

Degree is a limited measure of centrality in that it only identifies direct connections. By contrast, *k-path centrality* measures the extent to which an actor/vertex is indirectly connected to others, *betweenness centrality* measures the extent to which a vertex is an intermediary between otherwise disconnected groups of actors, and *eigenvector centrality* measures the extent to which a vertex is connected to highly-connected others [@RN4725]. Betweenness centrality tends to be relevant for empirical contexts (e.g., which manager gets credit for an idea) where an actor who occupies a position between groups can benefit by being a broker of information/ideas [@RN975; @RN4648]. By cont rast, this chapter is more concerned with whether a particular college tends to have indirect connections with some colleges but not others, and what is the subset of high schools this group of colleges tends to visit. Research question 2, motivated below, explores these ideas.

@@

This study engages with two, previously disconnected literatures. Scholarship on nonresident enrollment observes that many public research universities began growing nonresident enrollment following declines in state appropriations [@RN3753]. Given that nonresident tuition price is often two- to three-times resident price, it is unsurprising that nonresident students tend to be more affluent than resident students, are less likely to be first-generation students, and are less likely to be Black or Latinx [@RN3685]. Public research universities recruit nonresident students using institutional financial aid packages [@RN3762] and through recruiting interventions that target earlier stages in the enrollment funnel. Using nearly the same sample of 15 public research universities used in this study, @RN4758 found that 12 universities made more out-of-state visits than in-state visits and visits to out-of-state public high schools focused on affluent, predominantly white schools. Public research universities have a financial incentive to target private school students, who tend to come from affluent households able to pay nonresident tuition price without need-based aid. This study extends the literature on nonresident students by analyzing the scale of visits to private schools, the overlap and divergence in recruiting networks between public and private universities, and the characteristics of visited private schools.

We also engage with scholarship on linkages between private schools and private universities, a linkage that sociologists have observed for many decades. Private school students comprise a disproportionate share of enrollment at selective private universities [Author calculations based on HSLS:2009]. Practically, these linkages exist because private schools want to send their students to "good" universities while selective private universities want to enroll students who can pay private university tuition and who have desired academic profile and/or extracurricular activities [@RN3519]. Prior research has often focused on linkages between elite private schools and elite universities [@RN4407; @RN4935; @RN4640]. Whereas elite universities (e.g., Princeton, Stanford) experience strong student demand based on reputation, merely-selective private universities must incite demand by expending resources on recruiting. Prior research has not systematically analyzed connections between private schools and private universities that are selective, but not elite. Therefore, this study extends the literature by investigating private school visits for a sample of 14 selective private universities that exhibit variation in geographic location, religious affiliation, and ranking.

Our more important scholarly contribution is showing that these two literatures are both parts of a broader story about the changing nature of competition for students from privileged schools and communities. Selective private universities have been recruiting heavily from private schools for a long time [@RN4642; @RN3519; @RN4640]. By contrast, the pursuit of out-of-state private school students by public research universities is a relatively new phenomenon and represents competition in the traditional hunting grounds of selective private universities.

@@

However, public universities that made the greatest number of private school visits seemed to be targeting lower-achieving students at private schools targeted by particular private universities. For example, the University of Alabama visited 52.2% of the 625 private schools that Notre Dame visited, with most of the overlap at Catholic schools. Given the differences in academic prestige between them, the University of Alabama may not be competing for the same students as Notre Dame. Instead, they seem to be targeting Catholic school students of a similar economic echelon, who have lower academic achievement, and want to attend a football powerhouse. Additionally, CU-Boulder visited 42.4% of schools visited by Northwestern, which could be interpreted as CU-Boulder competing for lower-achieving students of a similar economic and cultural echelon at the schools targeted by Northwestern. Similarly, the University of South Carolina competes heavily at schools targeted by Emory, SMU, TCU, and Tulane.

@@

Market research suggests that off-campus recruiting visits will remain a vital means of connecting to students and guidance counselors, just as they were prior to the Covid pandemic. @RN4664 found that "travel" was the top-ranked top-ranked marketing/recruiting expenditure item for both public universities and private universities. During the height of Covid, off-campus recruiting visits ceased and were partially replaced by virtual counselor events [@parents_covid_rnl]. Admissions counselors are now traveling again. @RN4896 found that travel was the fifth ranked expenditure item for public universities and the second ranked expenditure item for private universities. When asked about areas of future expenditure growth, travel was the top ranked item for public universities and the second ranked item for private universities [@RN4896], suggesting that universities are increasing travel following the decline of the pandemic.

The @eab_nacac_counselor_2023 survey of high school counselors asked "how do you prefer to receive information from colleges?" High school visits by college representatives was by far the highest rated "channel" for private school counselors and was also the highest rated channel for public school counselors. By contrast, virtual visits by admissions representatives received low ratings by high school guidance counselors. Finally, @RN4895 asked universities to rank the most effective "outreach strategies for high school counselors." "Calling after visits" was deemed effective for 97% of private universities (the top strategy) and was deemed effective by 88% of public universities. What insights do we glean from these market research findings? In the era of mass marketing and data science, the pursuit of traditional college students from privileged schools and communities is still a hand-to-hand business and off-campus recruiting visits remain a vital means of recruiting these students.

Considering the practical importance of off-campus recruiting visits, what are the practical implications of our research? Prior scholarly and market research shows that students are sensitive to recruiting interventions, including recruiting visits [@RN4324; @RN4402]. Consequently, racial and socioeconomic composition of the schools that a university chooses to visit affects the racial and socioeconomic composition of the university. Our research on visits to public high schools [@RN4758] and this study on visits to private schools finds that our sample of public research universities and selective private universities tend to focus their recruiting efforts on affluent and predominantly white schools and communities. Some universities may publicly proclaim a commitment to social mobility and privately pursue an intentional recruiting strategy that targets affluent schools and communities. We offer practical guidance for universities that may be unintentionally implementing a recruiting strategy that privileges the privileged.
